2D Subquadratic Separable Distance Transformation for Path-Based Norms\thanks{This work has been mainly funded by ANR-11-BS02-009 and ANR-11-IDEX-0007-02 PALSE/2013/21 research grants.}

### 2014-04-18 latex/dgci2014/rebuttal.txt

        First of all, we would like to thank the reviewers for their
comments. In the following, We will comment some of major remarks
(please note that all remaining remarks, typos and suggestions will be
considered in the camera ready version if the paper is finally
accepted).

**Path based vs. Euclidean Metrics and quadratic complexity**

The first objective of the paper is to propose an algorithmic solution
to efficiently compute distance transformation for some path based
norms which mimics the parallel and separable approach used for
Euclidean and $L_p$ metrics. We will double-check the introduction to
make it clear that the paper focuses on that point.

Concerning the discussion about path based metric vs. Euclidean one,
we are convinced that path based norms are very interesting
mathematical objects and crucial in some applications (non-convex
domains, streaming DT, when the anisotropic nature of their unit ball
is consistent with the application, e.g. in crystallography...).

However, the huge literature about path-based metric optimization to
minimize the rotational dependency tends to illustrate that the
approximation quality (w.r.t. Euclidean distance) also matters. In this
context, the size of the mask is an important parameter (together with
optimized weights). Introduction will be rephrased to clarify this
context.

It is out of the scope of the paper (and won't be added in the
introduction) but please note that for medial axis extraction purposes
(defined as centers of maximal balls), linear  time algorithms
exists for the Euclidean metric [Coeurjolly-PAMI 2007, Hesselink PAMI
2008..]. For path based norms, pre-computations and look-up tables are
usually considered. As far as we know, it is still an active open
problem to prove that distances on shortest paths can be used to
extract the set of (or a subset of) maximal ball centers for the
metric [Strand, DGCI 2011].

**Experiments**

 - Graph size: we agree that graphs labels are quite small. We will
   correct this for the final version if the paper is accepted.

 - Scale-up w.r.t. to number of threads:  definitely
   doable. Experiments will be done and could be included in the final
   version (depending on the final version page limit).

 - Timings for low-size mask: zoom-in to the left-part of the graph
   can be added.

 - Comparison w.r.t. double-scan approach: doable too. However, we may
   end-up with issues with respect to page limit.

**Spelling, Grammar, Paper structure**

All apologizes for the bad English. The text will be double-checked
for the final version.

### 2014-05-13 latex/dgci2014/article.tex
raster scan

\begin{abstract}\sloppy
In many applications, separable algorithms have demonstrated their
efficiency to perform high performance volumetric computations, such
as distance transformation or medial axis extraction. In the
literature, several authors have discussed about the conditions on the
metric to be considered in a separable approach.  In this article, we
present generic separable algorithms to efficiently compute Voronoi
maps and distance transformations for a large class of
metrics. Focusing to path based norms (chamfer masks, neighborhood
sequences, ...), we detail a subquadratic algorithm to compute such
volumetric transformations in dimension 2. More precisely, we describe
a $O(\log^2{m}\cdot N^2)$ algorithm in dimension 2 for shapes in a
$N\times N$ domain with chamfer norm of size $m$.

@@

Since early works on digital geometry, distance transformation plays
an important role in many applications
\cite{Rosenfeld1966,Rosenfeld1968}. Given a finite input shape
$X\subset \Z^n$, the distance transformation labels each point in $X$
with the distance to its closest point in $\Z^n \setminus X$ (labeling
each point by the closest background point leads to Voronoi maps, see
below). Since such characterization is parametrized by a distance
function, many works have been done to address this distance
transformation problem with a trade-off between algorithmic
performances and the \emph{accuracy} of the digital distance function
with respect to the Euclidean one.  Hence, authors have considered
distances based on chamfer masks
\cite{Rosenfeld1968,borgefors,fouard:ivc:2005} or sequences of chamfer
masks \cite{Rosenfeld1966,mukherjee,Strand2008,DBLP:conf/dgci/NormandSE13};
the vector displacement based Euclidean distance
\cite{danielson,ragnemalm}; the Voronoi diagram based Euclidean
distance \cite{BreuEtAl95,Maurer2003} or the square of the Euclidean
distance \cite{Hirata,roerdnik}.  For Euclidean or $L_p$ metrics,
separable volumetric computations have demonstrated to be very
efficient: optimal $O(n\cdot N^n)$ time algorithms for shapes in $N^n$
domains, optimal multi-thread/GPU implementation\ldots (please refer
to \cite{dcoeurjo_ChapDTWADGMM} for a discussion). For path-based
metrics (chamfer mask, -weighted- neighborhood sequences,\ldots), two
main techniques exist to compute the distance transformation. The
first one considers a weighted graph formulation of the problem, and
shortest path in graphs using Dijkstra-like algorithm can be
considered. If $m$ denotes the size of the chamfer mask, computational
cost could be in $O(m\cdot N^n)$ using a cyclic bucket data structure
\cite{verwer_uniform}. Another approach consists in a raster scan of
the domain: first the chamfer mask is decomposed into two disjoint
sub-masks. Then the domain grid points are scanned in a given order
(consistent with the sub-masks construction) and a local minimum
computation is performed and then propagated
\cite{Rosenfeld1966,borgefors}. Scanning the domain twice for each
sub-masks leads to the distance transformation. Again, we end up with
a $O(m\cdot N^n)$ computational cost. This complexity is said to be
quadratic in $m$ and $N$ since for arbitrary large image $N$, even if
$m \ll N$, $m$ has to be in $O(N^n)$ to have accurate asymptotic DT.
Please note that Dijkstra's graph approach allows us to defined
constrained distance transformation (\emph{i.e.} geodesic metric),
both separable approaches and raster-scan for path-based metrics are
only dedicated to compact convex domains (usually hyper-rectangular)
distance transformation.

@@

framework for separable distance transformation and conditions on the
metric to be consistent with this model. Then, we describe
subquadratic algorithms in dimension 2 to compute error-free distance

@@

  F$.

@@

@@

characterize metric spaces $(\Z^n, \R , d_{L_p})$. Rounding the \sloppy
distance function $(\Z^n,\Z, \lceil d_{L_p}\rceil)$ is a metric
space \cite{klette_book}.

@@

$\Z^n\setminus$ and the lattice $\Z^n$. If a digital point $a$ belongs
to a Voronoi diagram $d-$facet ($1\leq d< n$), $a$ is equidistant to 2

@@

  vectors with non-empty vectors and containing at least a basis of

@@

construct chamfer masks approximating the Euclidean metric and being
induced by a norm. In the following, we focus on such \emph{chamfer
  norms}.  To evaluate distances between two digital points for a

@@

  Given a Chamfer mask $\M$ inducing a norm, the rational ball
  associated with $\M$ is the polytope

@@

Rational ball $\B$ can also be described in a H-representation of

@@

  of $A$.}$ $f$ being the number of row in $A$ and the number of

@@

Among path-based digital metric, we can also introduce (weighted)
neighborhood sequences
\cite{Rosenfeld1966,mukherjee,Strand2008,DBLP:conf/dgci/NormandSE13}
whose objective is to have better approximation of the Euclidean
metric from sequences of elementary chamfer masks.  A key result have
been demonstrated in \cite{DBLP:conf/dgci/NormandSE13} stating that
for such distance functions, a minimal parameter polytope
representation exists and that distances can be obtained from similar
expression to (\ref{eq:1}):

@@

In the following, we
consider Voronoi map approach as defined in \cite{BreuEtAl95}: Let us
first define hyper-rectangular image $I_X:
[1..N_1]\times\ldots\times[1..N_n] \rightarrow \{0,1\}$ such that $X\subset
I_X$ and $I_X(a)= 1$ for $a\in [1..N_1]\times\ldots\times[1..N_n]$ iff $a\in
X$ ($I_X(a)=0$ otherwise). In dimension 2, we first process each row
of the input image to create independent 1D Voronoi maps along the
first dimension for the metric. Then, for each further dimension, the
partial Voronoi map $\Pi_X$ is updated using one dimensional
independent processes on image spans along the $i^{th}$ dimension.

@@

  same coordinates except at $i$.} In this process, metric information
are embedded in the following two predicates (see

@@

\If{ i = 1}

@@

    \If{$I_X(s) = 1$}

@@

    \While{($k < |L_S|$) and not(\textsc{Closest}($s^i$, $L_S[k]$, $L_S[k+1]$))}

@@

  \caption{Voronoi map construction on 1D image span $S$ along the
    $i^{th}$ dimension.\label{alg:voro}}
\end{algorithm}
\begin{figure}
  \begin{center}
    \subfigure[]{ \begin{overpic}[width=2cm]{figs/hiddenByTrue}
    \put(26,5){$a$}
    \put(0,35){$b$}
    \put(25,75){$c$}
    \put(55,-15){$S$}
    \end{overpic}}
    \subfigure[]{\begin{overpic}[width=2cm]{figs/hiddenByFalse}
    \put(37,8){$a$}
    \put(23,37){$b$}
    \put(30,72){$c$}
    \put(55,-15){$S$}
    \end{overpic}}~~~~~~~~~~
    \subfigure[]{\begin{overpic}[width=1.1cm]{figs/closest}
    \put(-5,52){$a$}
    \put(2,78){$b$}
    \put(47,55){$c$}
    \put(38,-15){$S$}
      \end{overpic}
    }
  \end{center}
  \caption{\small Geometrical predicates for Voronoi map construction:
    \textsc{HiddenBy}($a,b,c,S)$ returns false in $(a)$ and true in
    $(b)$ (straight segments correspond to Voronoi diagram
    edges). $(c)$ illustrates the \textsc{Closest}($a,b,c$) predicate
    for $c\in S$.}
  \label{fig:predicates}
\end{figure}

In other words, \textsc{HiddenBy} returns true if and only if the
Voronoi cell of sites $a$ and $c$ \emph{hide} the Voronoi cell of $b$
along $S$.  For $L_1$, $L_2$ and $L_\infty$ metrics,
\textsc{Closest} and \textsc{HiddenBy} predicates can be computed in
$O(1)$ \cite{BreuEtAl95,Hirata,roerdnik}. Hence, Algorithm
\ref{alg:voro} is in $O(N_i)$ for dimension $i$, leading to an overall
computational time for the Voronoi map and distance transformation
problem in $\Theta(n\cdot N^n)$ (if we assume that $\forall i\in[1\ldots
  n], N_i=N$).

In \cite{Hirata} or \cite{Maurer2003}, authors discussed about
conditions on the metric $d$ to ensure that Algorithm \ref{alg:voro}
is correct. The key property can be informally describe as follows:
given two points $a,b\in\Z^n$ such that $a_i<b_i$ and a straight line
$l$ along the $i^{th}$ direction and if we denote by $v_l(a)$
(resp. $v_l(b)$) the intersection between the Voronoi cell of $a$
(resp. $b$) and l, then $v_l(a)$ and $v_l(b)$ are simply connected
Euclidean segments and $v_l(a)$ appears \emph{before} $v_l(b)$ on $l$
(so called \emph{monotonicity property} in \cite{Maurer2003} and is
related to \emph{quadrangle inequality} in \cite{Hirata}).  To sum
up these contributions, we have the following proposition:
\begin{proposition}[Metric conditions \cite{Hirata}]
\label{prop:voronoi-map-from-1}
  Let $d$ be a metric induced by a \textbf{norm whose unit ball is
    symmetric with respect to grid axes} and if distance comparison
  predicate is exact, Algorithm 1 is correct and returns a Voronoi map
  $\Pi_X$.
\end{proposition}
When implementing Algorithm 1, we say that the distance comparison
predicate is exact if we can compare two distances, \emph{e.g.}
\textsc{Closest} predicate, without error. Hence, Algorithm \ref{alg:voro}
provides exact Voronoi map computation for $L_p$ norms. In this case,
distance comparison predicate can be implemented error-free from
integer number comparison considering the $p$ power of the distance
function $\left (d_{L_p}\right)^p$.  Proposition
\ref{prop:voronoi-map-from-1} also implies that most chamfer norms and
neighborhood sequences based norms can also be considered in separable
Algorithm \ref{alg:voro} (see Fig.~\ref{fig:metrics}). We just need
algorithmic tools to efficiently implement both \textsc{Closest} and
\textsc{HiddenBy} predicates.
\begin{figure}[htbp]
 {\includegraphics[width=1.65cm]{image-ball-1}}
 {\includegraphics[width=1.65cm]{image-ball-1-4}}
 {\includegraphics[width=1.65cm]{image-ball-2}}
 {\includegraphics[width=1.65cm]{image-ball-4}}
 {\includegraphics[width=1.65cm]{image-ball-43-1}}
 {\includegraphics[width=1.65cm]{image-ball-chamf34}}
 {\includegraphics[width=1.65cm]{image-ball-chamf5711}}
  \caption{Balls for different metrics satisfying Proposition \ref{prop:voronoi-map-from-1}: \emph{(from
      left to right)} $L_1$, $L_{1.4}$, $L_2$, $L_4$, $L_{43.1}$,
    $\M_{3-4}$ and $\M_{5-7-11}$.}
  \label{fig:metrics}
\end{figure}
%
\vspace{-1cm}
\subsection{A First Generic Adapter}
\label{sec:first-gener-adapt}
We first detail the overall computational cost of Algorithm
\ref{alg:voro}:
\begin{lemma}
\label{lem:generic}
Let $(\Z^n, F, d)$ be a metric space induced by a norm with axis
symmetric unit ball. If $C$ denotes the computational cost of
\textsc{Closest} predicate and $H$ is the computational cost of the
\textsc{HiddenBy} predicate, then Algorithm \ref{alg:voro} is in
$O(n\cdot N^n\cdot (C+H))$.
\end{lemma}
From \cite{BreuEtAl95,roerdnik}, $C=H=O(1)$ for $L_1,L_2$ and $L_\infty$
norms.  For prescribed norm $d$, we first define generic
Algorithms~\ref{alg:genclosest}, \ref{alg:genedge} and
\ref{alg:genhidden}: from some evaluations of $d$, \textsc{HiddenBy}
predicate is obtained by a dichotomic search on the 1D image span $S$
to localize the abscissa of Voronoi edges of sites $\{a,b\}$ and
$\{b,c\}$ (see Fig.~\ref{fig:algo}).

@@

\begin{algorithm}[H]\footnotesize
  $v_{ab}$ = \textsc{VoronoiEdge}$(a,b,s^1, s^{N_i})$\;
  $v_{bc}$ = \textsc{VoronoiEdge}$(b,c,s^1, s^{N_i})$\;
  \Return $(v_{ab}> v_{bc})$\;
  \caption{\footnotesize Generic \textsc{HiddenBy}($a,b,c\in\Z^n; S$ in the $i^{th}$
    direction) $a_i<b_i<c_i$.\label{alg:genhidden}}
\end{algorithm}
\begin{lemma}
\label{lem:genericDich}
   Let $(\Z^n, F, d)$ be a metric space induced by a norm with axis
   symmetric unit ball,   from Algorithms \ref{alg:genclosest} and
   \ref{alg:genhidden}, we have $H=O(C\cdot \log{N})$.
\end{lemma}
\begin{proof}
  First, the generic \textsc{VoronoiEdge} is dichotomic in
  $O(\log{N})$ steps and $O(C)$ for each one. \textsc{VoronoiEdge} is
  in $O(C\cdot\log{N})$.  To prove the correctness of
  \textsc{VoronoiEdge} (and thus Alg.~\ref{alg:voro}), we use the
  convexity of the metric and the quadrangle property: since
  $a_i<b_i$, all grid points closer to $a$ than $b$ in $S$ (if exist)
  are \emph{lower} than all pixels on $S$ closer to $b$ than to $a$
  (if exist too). Thanks to test in line $4$, recursive call maintain
  this invariant. Note that tests $2-5$ handle the fact that the edge
  may not belong to $S$.$\Box$
\end{proof}
If we consider a chamfer norm with $f$ facet rational ball, Eq.~(\ref{eq:1})
suggests that $C=O(f)$. Hence, we have the following corollary:
\begin{corollary}
  Let $\M$ be a chamfer norm whose rational ball has $f$ facets,
  separable exact Voronoi map $\Pi_X$ can be obtained in
    $O(n\cdot N^n\cdot f \cdot \log{N})$.
\end{corollary}
Please remember that naive implementation of chamfer mask distance
transformation using raster scan approach would lead to a $O(f\cdot
N^n)$ computational cost. In the following sections, we use the convex
structure of $\B$ to design a first subquadratic algorithm for chamfer
norms.

@@

Let us consider a 2D chamfer norm $\M$ with $m$ weighted vectors (note
that $f\EqDef|\B| = m$ in 2D). We suppose that vectors
$\{\vec{v}^k\}_{k=1\ldots m}$ are sorted counterclockwise. We define a wedge
as a pair $(\vec{v}^k,\vec{v}^{k+1})$ of vectors. To each wedge is
associated a row $A_k$ in the minimal H-representation of $A$ ($A_k$
can also be seen as a  --non-unitary-- normal vectors to $\B$ facets
\cite{DBLP:journals/pr/NormandE09}). Using similar notations,
\cite{Thiel_hdr,Strand2008} demonstrate that the distance evaluation
of point $a$ can be obtained in two steps: First, we compute the wedge
$(\vec{v}^k,\vec{v}^{k+1})$, $a$ belongs to. Then,
\begin{equation}
\label{eq:2}
  d_\M(O,a) = A_k\cdot a^T
\end{equation}

@@

Please note that in practical implementations, we can use symmetries
in $\M$ vectors to only work on restrictions of chamfer mask
directions.  To optimize the \textsc{HiddenBy} predicate, we focus on
the \textsc{VoronoiEdge} function. Given two points $a$ and $b$
($a_i<b_i$) and a 1D image span $S$ along the $i^{th}$ dimension, we
have to find the lowest abscissa $e_i$ of the point $e$ on $S$ such
that $d(a,e) < d(b,e)$ and $d(a,e')\geq d(b, e')$ for any $e'$ with
$e'_i>e_i$.
Let us first suppose that we don't have $e$ but we know the two wedges
$(\vec{v}^{k},\vec{v}^{k+1})$ (resp. $(\vec{v}^j,\vec{v}^{j+1})$) such
that the vector $(e-a)^T$ (resp. $(e-b)^T$) belongs to (see
Fig.~\ref{fig:algo}). In this situation, we know that $e$ is the
solution of
\begin{equation}
\label{eq:3}
  A_k\cdot  (e-a)^T = A_j\cdot (e -b)^T\;.
\end{equation}
(since $e\in S$, we have one linear equation with only one unknown
$e_i$). As a consequence, if we know the two wedges the Voronoi edge
belongs to, we have the abscissa in $O(1)$ (see Algorithm
\ref{alg:chamedge} and Fig.~\ref{fig:algo}$-(c)$).
\begin{algorithm}[h]\footnotesize
  $(\vec{v}^k,\vec{v}^{k+1})$ = \textsc{VoronoiEdgeWedge}$(a,b,\vec{v}^1,\vec{v}^m, S)$\;
  $(\vec{v}^j,\vec{v}^{j+1})$ = \textsc{VoronoiEdgeWedge}$(b,a,\vec{v}^1,\vec{v}^m, S)$\;

>>>>

\begin{abstract}\sloppy
In many applications, separable algorithms have demonstrated their
efficiency to perform high performance and parallel volumetric
computations, such as distance transformation or medial axis
extraction. In the literature, several authors have discussed about
conditions on the metric to be considered in a separable approach.  In
this article, we present generic separable algorithms to efficiently
compute Voronoi maps and distance transformations for a large class of
metrics. Focusing to path based norms (chamfer masks, neighborhood
sequences, ...), we detail a subquadratic algorithm to compute such
volumetric transformation in dimension 2. More precisely, we describe
a $O(\log^2{m}\cdot N^2)$ algorithm for shapes in a $N\times N$ domain
with chamfer norm of size $m$.

@@

Since early works on digital geometry, distance transformation plays
an important role in many applications
\cite{Rosenfeld1966,Rosenfeld1968}. Given a finite input shape
$X\subset \Z^n$, the distance transformation labels each point in $X$
with the distance to its closest point in $\Z^n \setminus X$. Labeling
each point by the closest background point leads to Voronoi
maps. Since such characterization is parametrized by a distance
function, many authors address this distance transformation problem
with trade-offs between algorithmic performances and the
\emph{accuracy} of the digital distance function with respect to the
Euclidean one.  Hence, authors have considered distances based on
chamfer masks \cite{Rosenfeld1968,borgefors,fouard:ivc:2005} or
sequences of chamfer masks
\cite{Rosenfeld1966,mukherjee,Strand2008,DBLP:conf/dgci/NormandSE13};
the vector displacement based Euclidean distance
\cite{danielson,ragnemalm}; Voronoi diagram based Euclidean distance
\cite{BreuEtAl95,Maurer2003} or the square of the Euclidean distance
\cite{Hirata,roerdnik}.  For the Euclidean metric, separable
volumetric computations have demonstrated to be very efficient:
optimal $O(n\cdot N^n)$ time algorithms for shapes in $N^n$ domains,
optimal multi-thread/GPU implementation\ldots (please refer to
\cite{dcoeurjo_ChapDTWADGMM} for a discussion). For path-based metrics
(chamfer mask, -weighted- neighborhood sequences,\ldots), two main
techniques exist to compute the distance transformation. The first one
considers a weighted graph formulation of the problem and
Dijkstra-like algorithms on weighted graphs to compute distances. If
$m$ denotes the size of the chamfer mask, computational cost could be
in $O(m\cdot N^n)$ using a cyclic bucket data structure
\cite{verwer_uniform}. Another approach consists in a raster scan of
the domain: first the chamfer mask is decomposed into disjoint
sub-masks. Then the domain grid points are scanned in a given order
(consistent with the sub-mask construction) and a local computation is
performed before being propagated
\cite{Rosenfeld1966,borgefors}. Scanning the domain several times (one
per sub-mask) leads to the distance transformation values. Again, we
end up with a $O(m\cdot N^n)$ computational cost.

Beside specific applications which use the anisotropic nature of the
chamfer mask, rotational dependency is usually ensured with increasing
mask size $m$ together with weight optimization. In this context and
for arbitrarily large image $N$, both Dijsktra-like and raster scan
approach have computational cost which is quadratic with respect to
$N$ and $m$. In practical situations, $m \ll N$ but $m$ still needs to
be in $O(N^n)$ to have accurate asymptotic DT.

Please note that Dijkstra's graph approach allows us to defined
constrained distance transformation (\emph{i.e.} geodesic metric),
both separable approaches and raster-scan for path-based metrics are
only dedicated to compact convex domains (usually hyper-rectangular)
distance transformation.

@@

framework for separable distance transformation and metric conditions
to be consistent with this model. Then, we describe subquadratic and
parallel algorithms in dimension 2 to compute error-free distance

@@

  F$ (with $E$ such that for $a,b\in E$, $(b-a)\in EV$).

@@

\emph{digital}

@@

defines metric spaces $(\Z^n, \R , d_{L_p})$ which are not
\emph{digital}. However, rounding up the \sloppy distance function
$(\Z^n,\Z, \lceil d_{L_p}\rceil)$ is a digital metric space
\cite{klette_book}.

@@

$\Z^n\setminus X$ and the lattice $\Z^n$. If a digital point $a$ belongs
to a Voronoi diagram $d-$facet ($0\leq d< n$), $a$ is equidistant to 2

@@

  vectors with no null vectors and containing at least a basis of

@@

construct chamfer masks approximating the Euclidean metric. In the following, we focus on such \emph{chamfer
  norms} which are chamfer metric induced by a norm).  To evaluate distances between two digital points for a

@@

  Given a Chamfer norm $\M$, the
  rational ball associated with $\M$ is the polytope

@@

Rational ball $\B$ can also be described as the H-representation of

@@

  of $A$.}$ $f$ being the number of rows in $A$ and the number of

@@

Among path-based digital metric, (weighted) neighborhood sequences
have been proposed to have better approximation of the Euclidean
metric from sequences of elementary chamfer masks
\cite{Rosenfeld1966,mukherjee,Strand2008,DBLP:conf/dgci/NormandSE13}.
A key result have been demonstrated in
\cite{DBLP:conf/dgci/NormandSE13} stating that for such distance
functions, a minimal parameter polytope representation exists and that
distances can be obtained from similar expression to (\ref{eq:1}):

@@

In the following, we consider Voronoi map approach as defined in
\cite{BreuEtAl95}: Let us first define hyper-rectangular image $I_X:
     [1..N_1]\times\ldots\times[1..N_n] \rightarrow \{0,1\}$ such that
     $I_X(a)= 1$ for $a\in [1..N_1]\times\ldots\times[1..N_n]$ iff
     $a\in X$ ($I_X(a)=0$ otherwise). In dimension 2, each row of the
     input image are processed to create independent 1D Voronoi maps
     along the first dimension for the metric. Then, for each further
     dimension, the partial Voronoi map $\Pi_X$ is updated using one
     dimensional independent processes on image spans along the
     $i^{th}$ dimension.

@@

  same coordinates except at $i$.}. In this process, metric information
are embedded in the following  predicates (see

@@

\If{ $i == 1$ \tcp*[r]{Special case for the first  dimension}}

@@

    \If{$I_X(s) == 1$}

@@

    \While{($k < |L_S|$) and not(\textsc{Closest}($s$, $L_S[k]$, $L_S[k+1]$))}

@@

  \caption{Voronoi map construction on 1D image span $S$ along the
    $i^{th}$ dimension.\label{alg:voro}}
\end{algorithm}
\begin{figure}
  \begin{center}
    \subfigure[]{ \begin{overpic}[width=2cm]{figs/hiddenByTrue}
    \put(26,5){$a$}
    \put(0,35){$b$}
    \put(25,75){$c$}
    \put(55,-15){$S$}
    \end{overpic}}
    \subfigure[]{\begin{overpic}[width=2cm]{figs/hiddenByFalse}
    \put(37,8){$a$}
    \put(23,37){$b$}
    \put(30,72){$c$}
    \put(55,-15){$S$}
    \end{overpic}}~~~~~~~~~~
    \subfigure[]{\begin{overpic}[width=1.1cm]{figs/closest}
    \put(-5,52){$b$}
    \put(2,78){$c$}
    \put(47,55){$a$}
    \put(38,-15){$S$}
      \end{overpic}
    }
  \end{center}
  \caption{\small Geometrical predicates for Voronoi map construction:
    \textsc{HiddenBy}($a,b,c,S)$ returns false in $(a)$ and true in
    $(b)$ (straight segments correspond to Voronoi diagram
    edges). $(c)$ illustrates the \textsc{Closest}($a,b,c$) predicate
    for $c\in S$.}
  \label{fig:predicates}
\end{figure}

In other words, \textsc{HiddenBy} returns true if and only if the
Voronoi cells of sites $a$ and $c$ \emph{hide} the Voronoi cell of $b$
along $S$.  For $L_1$, $L_2$ and $L_\infty$ metrics,
\textsc{Closest} and \textsc{HiddenBy} predicates can be computed in
$O(1)$ \cite{BreuEtAl95,Hirata,roerdnik}. Hence, Algorithm
\ref{alg:voro} is in $O(N_i)$ for dimension $i$, leading to an overall
computational time for the Voronoi map and distance transformation
problem in $\Theta(n\cdot N^n)$ (if we assume that $\forall i\in[1\ldots
  n], N_i=N$).

In \cite{Hirata} or \cite{Maurer2003}, authors discussed about
conditions on the metric $d$ to ensure that Algorithm \ref{alg:voro}
is correct. The key property can be informally describe as follows:
given two points $a,b\in\Z^n$ such that $a_i<b_i$ and a straight line
$l$ along the $i^{th}$ direction and if we denote by $v_l(a)$
(resp. $v_l(b)$) the intersection between the Voronoi cell of $a$
(resp. $b$) and $l$, then $v_l(a)$ and $v_l(b)$ are simply connected
Euclidean segments and $v_l(a)$ appears \emph{before} $v_l(b)$ on $l$
(so called \emph{monotonicity property} in \cite{Maurer2003} and is
related to \emph{quadrangle inequality} in \cite{Hirata}).  To sum
up these contributions, we have the following sufficient conditions on
the metric:
\begin{proposition}[Metric conditions \cite{Hirata}]
\label{prop:voronoi-map-from-1}
  Let $d$ be a metric induced by a \textbf{norm whose unit ball is
    symmetric with respect to grid axes} and if distance comparison
  predicate is exact, Algorithm 1 is correct and returns a Voronoi map
  $\Pi_X$.
\end{proposition}
When implementing Algorithm 1, the distance comparison
predicate is exact if we can compare two distances, \emph{e.g.}
\textsc{Closest} predicate, without error.

For $L_p$ norms Algorithm \ref{alg:voro} provides exact Voronoi map
computation. Indeed, distance comparison predicate can be implemented
error-free from integer number comparisons considering the $p$ power of
the distance function $\left (d_{L_p}\right)^p$.

Proposition
\ref{prop:voronoi-map-from-1} also implies that most chamfer norms and
neighborhood sequence based norms can also be considered in separable
Algorithm \ref{alg:voro} (see Fig.~\ref{fig:metrics}). We just need
algorithmic tools to efficiently implement both \textsc{Closest} and
\textsc{HiddenBy} predicates.
\begin{figure}[htbp]
 {\includegraphics[width=1.65cm]{image-ball-1}}
 {\includegraphics[width=1.65cm]{image-ball-1-4}}
 {\includegraphics[width=1.65cm]{image-ball-2}}
 {\includegraphics[width=1.65cm]{image-ball-4}}
 {\includegraphics[width=1.65cm]{image-ball-43-1}}
 {\includegraphics[width=1.65cm]{image-ball-chamf34}}
 {\includegraphics[width=1.65cm]{image-ball-chamf5711}}
  \caption{Balls for different metrics satisfying Proposition \ref{prop:voronoi-map-from-1}: \emph{(from
      left to right)} $L_1$, $L_{1.4}$, $L_2$, $L_4$, $L_{43.1}$,
    $\M_{3-4}$ and $\M_{5-7-11}$.}
  \label{fig:metrics}
\end{figure}
%
\vspace{-1cm}
\subsection{A First Generic Adapter}
\label{sec:first-gener-adapt}
We first detail the overall computational cost of Algorithm
\ref{alg:voro}:
\begin{lemma}
\label{lem:generic}
Let $(\Z^n, F, d)$ be a metric space induced by a norm with axis
symmetric unit ball. If $C$ denotes the computational cost of
\textsc{Closest} predicate and $H$ is the computational cost of the
\textsc{HiddenBy} predicate, then Algorithm \ref{alg:voro} is in
$O(n\cdot N^n\cdot (C+H))$.
\end{lemma}
From \cite{BreuEtAl95,roerdnik}, $C=H=O(1)$ for $L_1,L_2$ and $L_\infty$
norms.  For prescribed norm $d$, we first define generic
Algorithms~\ref{alg:genclosest}, \ref{alg:genedge} and
\ref{alg:genhidden}: From some evaluations of $d$, \textsc{HiddenBy}
predicate is obtained by a dichotomic search on the 1D image span $S$
to localize the abscissa of Voronoi edges of sites $\{a,b\}$ and
$\{b,c\}$ (see Fig.~\ref{fig:algo}).

@@

\begin{algorithm}[H]\footnotesize
  $v_{ab}$ = \textsc{VoronoiEdge}$(a,b,s^1, s^{N_i})$\;
  $v_{bc}$ = \textsc{VoronoiEdge}$(b,c,s^1, s^{N_i})$\;
  \Return $(v_{ab}> v_{bc})$\;
  \caption{\footnotesize Generic \textsc{HiddenBy}($a,b,c\in\Z^n; S$ in the $i^{th}$
    direction) $a_i<b_i<c_i$.\label{alg:genhidden}}
\end{algorithm}
\begin{lemma}
\label{lem:genericDich}
   Let $(\Z^n, F, d)$ be a metric space induced by a norm with axis
   symmetric unit ball,   from Algorithms \ref{alg:genclosest} and
   \ref{alg:genhidden}, we have $H=O(C\cdot \log{N})$.
\end{lemma}
\begin{proof}
  First, the generic \textsc{VoronoiEdge} is dichotomic with
  $O(\log{N})$ steps and $O(C)$ for each one. \textsc{VoronoiEdge} is
  in $O(C\cdot\log{N})$.  To prove the correctness of
  \textsc{VoronoiEdge} (and thus Alg.~\ref{alg:voro}), we use the
  convexity of the metric and the quadrangle property: since
  $a_i<b_i$, all grid points closer to $a$ than $b$ in $S$ (if exist)
  are \emph{lower} than all pixels on $S$ closer to $b$ than to $a$
  (if exist too). Thanks to test in line $4$, recursive call maintain
  this invariant. Note that tests $2-5$ handle the fact that the edge
  may not belong to $S$.$\Box$
\end{proof}
If we consider a chamfer norm with $f$ facet rational ball, Eq.~(\ref{eq:1})
suggests that $C=O(f)$. Hence, we have the following corollary:
\begin{corollary}
  Let $\M$ be a chamfer norm whose rational ball has $f$ facets,
  separable exact Voronoi map $\Pi_X$ can be obtained in
    $O(n\cdot N^n\cdot f \cdot \log{N})$.
\end{corollary}
Please remember that naive implementation of chamfer mask distance
transformation using raster scan approach would lead to a $O(f\cdot
N^n)$ computational cost. In the following sections, we use the convex
structure of $\B$ to design a parallel subquadratic algorithm for chamfer
norms.

@@

Let us consider a 2D chamfer norm $\M$ with $m$ weighted vectors (note
that $f\EqDef|\B| = m$ in 2D). We suppose that vectors
$\{\vec{v}^k\}_{k=1\ldots m}$ are sorted counterclockwise. We define a wedge
as a pair $(\vec{v}^k,\vec{v}^{k+1})$ of vectors. To each wedge is
associated a row $A_k$ in the minimal H-representation of $A$ ($A_k$
can also be seen as a  --non-unitary-- normal vector to $\B$ facets
\cite{DBLP:journals/pr/NormandE09}). Using similar notations,
\cite{Thiel_hdr,Strand2008} demonstrate that the distance evaluation
of point $a$ can be obtained in two steps: First, we compute the wedge
$(\vec{v}^k,\vec{v}^{k+1})$, $a$ belongs to. Then,
\begin{equation}
\label{eq:2}
  d_\M(O,a) = A_k\cdot a^T
\end{equation}

@@

Please note that in practical implementations, we can use symmetries
in $\M$  to only work on restrictions of chamfer mask
directions, so called \emph{generator} in the literature.  To optimize the \textsc{HiddenBy} predicate, we focus on
the \textsc{VoronoiEdge} function. Given two points $a$ and $b$
($a_i<b_i$) and a 1D image span $S$ along the $i^{th}$ dimension, we
have to find the lowest abscissa $e_i$ of the point $e$ on $S$ such
that $d(a,e) < d(b,e)$ and $d(a,e')\geq d(b, e')$ for any $e'$ with
$e'_i>e_i$.
Let us first suppose that we do not have $e$ but we know the two wedges
$(\vec{v}^{k},\vec{v}^{k+1})$ (resp. $(\vec{v}^j,\vec{v}^{j+1})$) such
that the vector $(e-a)^T$ (resp. $(e-b)^T$) belongs to (see
Fig.~\ref{fig:algo}). In this situation, we know that $e$ is the
solution of
\begin{equation}
\label{eq:3}
  A_k\cdot  (e-a)^T = A_j\cdot (e -b)^T\;.
\end{equation}
(since $e\in S$, we have one linear equation with only one unknown
$e_i$). As a consequence, if we know the two wedges the Voronoi edge
belongs to, we have the abscissa in $O(1)$ (see Algorithm
\ref{alg:chamedge} and Fig.~\ref{fig:algo}$-(c)$).
\begin{algorithm}[h]\footnotesize
  $(\vec{v}^k,\vec{v}^{k+1})$ = \textsc{VoronoiEdgeWedge}$(a,b,\vec{v}^1,\vec{v}^m, S)$\;
  $(\vec{v}^j,\vec{v}^{j+1})$ = \textsc{VoronoiEdgeWedge}$(b,a,\vec{v}^1,\vec{v}^m, S)$\;
